Approximation as Differentiation: Notes

### 2025-05-16 main/formal-structures.tex
Prefer direct composition notation for reindexing functor.

This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
between them as morphisms. The Grothendieck construction $\Grothendieck{}T$ is the ``total category'' which
gloms all such $\Set$-indexed families of bounded lattices into the same category, with reindexing morphisms
transporting between data indexed by different sets.

\begin{definition}[Functor category of $X$-indexed families]
\label{def:fam:Fam}
For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
\end{definition}

@@

\begin{definition}[Reindexing]
For any function $f: X \to Y$ define the \emph{reindexing} functor $\reindex{-}{f}: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family
$\reindex{F}{f} = F \comp f$, regarding $f$ as a functor between $X$ and $Y$ as discrete categories.
\end{definition}

@@

Of interest to us is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
category which sends any set $X$ to the functor category $\Func{X}{\cat{C}}$ and any function $f: X \to Y$ to
the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$.

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \reindex{\partial Y}{f}$ in $\Fam(X,\cat{C})$.

>>>>

\subsubsection{Grothendieck construction for indexed categories}

This pattern of objects and morphisms is exactly the \emph{Grothendieck construction} for a particular indexed
category $T: \Set^\op \to \Cat$, namely the functor which assigns to every set $X$ the functor category $T(X)
= \Func{X}{\LatGal}$. Here the (discrete) category $X$ is a plain set and so $\Func{X}{\LatGal}$ is exactly
the category of $X$-indexed families of bounded lattices, with $X$-indexed families of Galois connections
between them as morphisms. The Grothendieck construction $\Grothendieck{}T$
(\defref{grothendieck:Grothendieck} below) is the category of all $\Set$-indexed families of bounded lattices,
together with morphisms that account for both structure within each family and how families transform under
reindexing.

First we formalise $X$-indexed families and their categories.

\begin{definition}[Category of $X$-indexed families]
\label{def:fam:Fam}
For any set $X$ (interpreted as a discrete category) and any category $\cat{C}$, recall that the functor
category $\Func{X}{\cat{C}}$ has as objects all functors $F: X \to \cat{C}$, namely $X$-indexed families of
objects of $\cat{C}$, and as morphisms from $F \to G$, all families of morphisms $\eta(x): F(x) \to G(x)$ in
$\cat{C}$ for every $x \in X$, with naturality trivial because $X$ only has identity morphisms.
\end{definition}

@@

Any function $f: X \to Y$ induces a \emph{reindexing} functor $- \comp f: \Func{Y}{\cat{C}} \to
\Func{X}{\cat{C}}$ which sends any $Y$-indexed family $F$ over $\cat{C}$ to the $X$-indexed family $F \comp f$
(regarding $f$ as a functor between $X$ and $Y$ as discrete categories) and any family of morphisms $\eta: F
\to G$ to $\eta \comp f: F \comp f \to G \comp f$ with $(\eta \comp f)_x = \eta_{f(x)}$.

@@

Of interest here is the Grothendieck construction for $\Func{-}{\cat{C}}: \Set^{\op} \to \Cat$, the indexed
category which sends any set $X$ to the category of $X$-indexed families $\Func{X}{\cat{C}}$ and any function
$f: X \to Y$ to the reindexing functor $- \comp f: \Func{Y}{\cat{C}} \to \Func{X}{\cat{C}}$ .

@@

\noindent In $\Fam(\cat{C})$ the objects $(X, \partial X)$ are thus sets $X$ paired with indexed families
$\partial X: X \to \cat{C}$ and morphisms $(f, \partial f): (X, \partial X) \to (Y, \partial Y)$ are functions
$f: X \to Y$ paired with morphisms $\partial f: \partial X \to \partial Y \comp f$ in $\Func{X}{\cat{C}}$.
